

\let\textcircled=\pgftextcircled\chapter{Results and Validation}

Introduction
Briefly summarize the goals of the RL experiments.
Outline the structure of the results chapter.

\section{Final Experimental Setup}
Experimental Setup
Detail the final setup of the training environment in Unity.
Describe the neural network architecture and the parameters used.
Outline the training procedure, including any pre-training steps.

\section{Training Results}
Training Results
Present the learning curves for the agent, showing reward over time.
Include a table or graph of the agent’s performance metrics at various checkpoints.
Discuss any unexpected behaviors or anomalies observed during training.

\section{Evaluation Of Agent Performance}
Evaluation of Agent Performance
Explain the methods used to evaluate the trained agent (e.g., test episodes in varied conditions).
Present the results of these evaluations in a systematic manner (e.g., tables, graphs).
Discuss the agent’s ability to generalize from its training to new scenarios.

Comparison with Baselines
If there are baseline models or industry standards, compare your results against these.
Use statistical methods to determine the significance of the results.

Ablation Studies
Discuss any ablation studies conducted (if any) to understand the contribution of different components of the system.
Present the impact of removing/modifying certain parts of the agent or the environment.

Visualizations and Case Studies
Provide visual representations of agent behavior, such as plots of trajectories or state visitation frequencies.
Include screenshots or video stills from Unity to illustrate successful and unsuccessful episodes.

Summary
Sum up the main findings from the results.
Discuss any limitations of these results or the experimental setup.


\section{Evaluation Introduction}
Evaluation
Introduction
Reflect on the purpose of the evaluation.
Outline what the chapter will cover.

Methodology for Evaluation
Explain the metrics used to evaluate the agent’s performance.
Describe the process of collecting evaluation data.

Performance Analysis
Analyze the agent’s performance in depth, possibly breaking down by different types of tasks or challenges it faced.
Use statistical methods to discuss the performance variability.

Robustness and Generalization
Evaluate the agent's robustness to changes in the environment (e.g., different wind conditions, system perturbations).
Discuss the agent's ability to generalize from its training environment to similar but unseen environments.

Comparison to Human Performance
If relevant, compare the agent’s performance to that of a human completing the same task.
Discuss any qualitative differences in approaches to the task between the agent and humans.

Discussion of the Evaluation Results
Interpret the results in the context of the project's objectives.
Discuss the strengths and weaknesses of the agent as revealed by the evaluation.

Implications for Future Work
Suggest how the results might inform future projects or the development of RL agents for similar tasks.
Discuss any additional experiments or data that would be valuable.

Conclusion
Summarize the main takeaways from the evaluation.
Discuss the conclusions that can be drawn about the RL agent's performance and the potential for real-world application.